---
title: 'dplr: Practice'
author: "R Workshop"
output:
  html_document:
    css: ../workshop.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(warning=FALSE, message=FALSE)
```


### Earthquake data

```{r load-quakes, echo=FALSE}
data(quakes)
require(dplyr)
```

1.   Show a command you could use to find the average (mean) magnitude of all earthquakes that occurred below the median depth of all earthquakes.

```{r}
quakes %>%
  filter(depth > median(quakes$depth)) %>%
  summarize(mean(mag))
```


1.   How does the average magnitude of these "deep" earthquakes compare to the magnitude of "shallow" (above median depth) earthquakes?

```{r}
quakes %>%
  filter(depth < median(quakes$depth)) %>%
  summarize(mean(mag))
```


1.   Show the command you would use to get the row(s) with the smallest value in the `depth` column.

```{r}
quakes %>%
  filter(depth == min(quakes$depth))
```


1.   Show the command you would use to create a new data frame named `quakes_sorted` that stores the data in sorted order by `depth`.

```{r}
quakes_sorted <- quakes %>%
  arrange(depth)

quakes_sorted
```


* * *

### Flights data

```{r load-flights, eval = FALSE}
library(pnwflights14)
data(flights)
```

1. Say I fly very often between PDX airport in Portland and Logan Airport in Boston. Based on the 2014 data, which day of the week would you recommend that I buy a ticket for if I want to minimize delays?

```{r}
# days_of_week <- c("Wednesday", "Thursday", "Friday", "Saturday", "Sunday", "Monday", "Tuesday")
# days_of_week_2014 <- rep(days_of_week, len=365)
# 
# flights_dow <- flights %>%
#   
# 
# flights_doy <- flights %>%
#   arrange(year, month, day) %>%
#   group_by(year, month, day) %>%
#   summarize(avg_delay = mean(dep_delay, na.rm = TRUE))
# 
# flights_doy$dow <- days_of_week_2014
#   
# flights_dow <- flights_doy %>%
#   group_by(dow) %>%
#   summarize(avg_delay = mean(avg_delay))
# 
# flights_dow$dow[which.min(flights_dow$avg_delay)]

get_day_of_week <- function(d, m, y) {
  weekdays(as.Date(paste(d, "-", m, "-", y, sep = ""), '%d-%m-%Y'))
}

pdx_bos <- flights %>%
  filter((origin == "PDX" & dest == "BOS") | (origin == "BOS" & dest == "PDX")) %>%
  mutate(dow = get_day_of_week(day, month, year)) %>%
  group_by(dow) %>%
  summarize(avg_delay = mean(dep_delay, na.rm = TRUE), count = n())

pdx_bos
```

```{r}

flights_by_time <- flights %>%
  mutate(scheduled = (dep_time - dep_delay) %% 2400 )

plot(flights_by_time$scheduled, flights_by_time$dep_delay)

```


1. What advice do you have for the best airline for this route?


* * *

### Back to Global Warming

Go back to the vector `tData` that you used for the project. Use this data and
maybe `dplyr` to form a dataframe that will make the analysis much more straightforward.

```{r}
load(url('http://www.stolaf.edu/people/olaf/cs125/project1.RData'))
```

```{r}
faren_data <- (tData/100 + 14)*1.8 + 32

data_matrix <- matrix(data = faren_data, nrow = length(faren_data)/12, ncol = 12, byrow = TRUE)

# Set column names
colnames(data_matrix) <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
rownames(data_matrix) <- 1881:2012

tdata.df <- as.data.frame(as.table(data_matrix), stringsAsFactors = FALSE)

temp_data <- tdata.df %>%
  transmute(Temperature = Freq, Month = as.factor(Var2), Year = as.numeric(Var1)) %>%
  arrange(Year)
```


